Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-10-21 clustering_infinite_model.tex
fixed errors in bib file and corresponding references in the LaTeX files

The hypergeometric U-function 
(also called Tricomi's confluent hypergeometric function) has the integral representation 
$$U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt$$ which holds for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $Re(a), Re(z) >0$, see \cite[p.255]{Erdelyi1953}. Applying the change of variables 
$t=\frac{1-s}{s}$ (i.e. $dt = -s^{-2} ds$ and $s = \frac{1}{t+1}$) yields the integral 
$\int_1^0 e^{-z(-1+\frac{1}{s})} (\frac{1-s}{s})^{a-1} (\frac{1}{s})^{b-a-1} (-s^2)ds = e^z\int_0^1 e^{-z/s} s^{-b}(1-s)^{a-1} ds$, i.e.
\begin{align*}
U(a,b,z) = \frac{e^z}{\Gamma(a)} \int_0^1 s^{-b} (1-s)^{a-1} e^{-z/s} ds
\end{align*}
Plugging in $a=2\alpha+1 >0$, $b=-2\alpha+k+1$, $z=\xi>0$, gives
\begin{align*}
\int_0^1 z_0^{2\alpha-k-1} e^{-\xi/z_0} (1-z_0)^{2\alpha} dz_0 = \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi)
\end{align*}

Let $a, c \in \R$, $\xi, b >0$ positive real numbers. 
Using the integral definition of the incomplete B-function, the change of variables $s=1-t$ gives:
\begin{align*}
\int_0^1 z^a e^{-\xi/z} B^-(1-z;b,c) dz &=\int_0^1 z^a e^{-\xi/z} \int_0^{1-z} t^{b-1} (1-t)^{c-1} dt dz \\
&=\int_0^1 z^a e^{-\xi/z} \int_z^1 s^{c-1} (1-s)^{b-1} ds dz
\end{align*}
Then changing the order of integration and using the substitution $u = \xi/z$  and recognizing the upper incomplete $\Gamma$-function yields
\begin{align*}
&=\int_0^1 \int_0^s z^a e^{-\xi/z} dz s^{c-1} (1-s)^{b-1} ds \\
&=\int_0^1 \int_{\xi/s}^\infty \xi^{a+1} u^{-a-2} e^{-u} du s^{c-1} (1-s)^{b-1} ds \\
&= \xi^{a+1} \int_0^1 \Gamma^+(-a-1,\xi/s) s^{c-1} (1-s)^{b-1} ds
\end{align*}
Now, the incomplete $\Gamma$-function has a representation in terms of Meijer's $G$-function $$\Gamma^+(-a-1,\xi/s) = \MeijerG{2}{0}{1}{2}{1}{-a-1,0}{\frac{\xi}{s}}$$ which holds for any $a\in \R$ and $s>0$ (see appendix; note that for a fixed second argument, the upper incomplete $\Gamma$-function is entire in the first argument, see \cite[pp. 899, 1032ff.]{Gradshteyn2007}).  Therefore, we get the integral

%formula 8.351.3, p. 
 %(see the wolfram functions website \url{http://functions.wolfram.com/06.06.26.0005.01}, it can also be found in the paper \url{https://arxiv.org/pdf/1507.05571}; , so we get the integral
\begin{align*}
= \xi^{a+1} \int_0^1 s^{c-1} (1-s)^{b-1} \MeijerG{2}{0}{1}{2}{1}{-a-1,0}{\frac{\xi}{s}}ds
\end{align*}
applying the inversion identity for Meijer's $G$-function (see \cite[p. 209, 5.3.1.(9))]{Erdelyi1953})
\begin{align*}
=\xi^{a+1} \int_0^1 s^{c-1} (1-s)^{b-1} \MeijerG{0}{2}{2}{1}{2+a,1}{0}{\frac{s}{\xi}}ds
\end{align*}
This is the Euler transform of Meijer's $G$-function (see \cite[p. 214, 5.5.2.(5)]{Erdelyi1953}) and (as the conditions $2+1<2(0+2)$ and $|\arg(\xi^{-1})| < \frac{\pi}{2}$ (as $\xi>0$) and $1-c-b<1-c$ (as $b>0$)  are satisfied) equals
\begin{align*}
=\xi^{a+1} \Gamma(b) \MeijerG{0}{3}{3}{2}{1-c,2+a,1}{0,1-c-b}{\xi^{-1}}
\end{align*}
Using again the inversion identity for Meijer's $G$-function
\begin{align*}
=\xi^{a+1} \Gamma(b) \MeijerG{3}{0}{2}{3}{1,b+c}{c,-1-a,0}{\xi}
\end{align*}
Plugging in $a= 6\alpha-k-3$, $b=2\alpha$, $c=3-4\alpha$ results in
\begin{align*}
=\xi^{6a-k-2} \Gamma(2\alpha) \MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}
\end{align*}
%(note that this agrees with the result in the old approach, after `bringing the power of $\xi$ into Meijer's $G$-function'.)

>>>>

The hypergeometric U-function 
(also called Tricomi's confluent hypergeometric function) has the integral representation 
$$U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt$$ which holds for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $Re(a), Re(z) >0$, see \cite[p.255]{erdelyi1953higher}. Applying the change of variables 
$t=\frac{1-s}{s}$ (i.e. $dt = -s^{-2} ds$ and $s = \frac{1}{t+1}$) yields the integral 
$\int_1^0 e^{-z(-1+\frac{1}{s})} (\frac{1-s}{s})^{a-1} (\frac{1}{s})^{b-a-1} (-s^2)ds = e^z\int_0^1 e^{-z/s} s^{-b}(1-s)^{a-1} ds$, i.e.
\begin{align*}
U(a,b,z) = \frac{e^z}{\Gamma(a)} \int_0^1 s^{-b} (1-s)^{a-1} e^{-z/s} ds
\end{align*}
Plugging in $a=2\alpha+1 >0$, $b=-2\alpha+k+1$, $z=\xi>0$, gives
\begin{align*}
\int_0^1 z_0^{2\alpha-k-1} e^{-\xi/z_0} (1-z_0)^{2\alpha} dz_0 = \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi)
\end{align*}

Let $a, c \in \R$, $\xi, b >0$ positive real numbers. 
Using the integral definition of the incomplete B-function, the change of variables $s=1-t$ gives:
\begin{align*}
\int_0^1 z^a e^{-\xi/z} B^-(1-z;b,c) dz &=\int_0^1 z^a e^{-\xi/z} \int_0^{1-z} t^{b-1} (1-t)^{c-1} dt dz \\
&=\int_0^1 z^a e^{-\xi/z} \int_z^1 s^{c-1} (1-s)^{b-1} ds dz
\end{align*}
Then changing the order of integration and using the substitution $u = \xi/z$  and recognizing the upper incomplete $\Gamma$-function yields
\begin{align*}
&=\int_0^1 \int_0^s z^a e^{-\xi/z} dz s^{c-1} (1-s)^{b-1} ds \\
&=\int_0^1 \int_{\xi/s}^\infty \xi^{a+1} u^{-a-2} e^{-u} du s^{c-1} (1-s)^{b-1} ds \\
&= \xi^{a+1} \int_0^1 \Gamma^+(-a-1,\xi/s) s^{c-1} (1-s)^{b-1} ds
\end{align*}
Now, the incomplete $\Gamma$-function has a representation in terms of Meijer's $G$-function $$\Gamma^+(-a-1,\xi/s) = \MeijerG{2}{0}{1}{2}{1}{-a-1,0}{\frac{\xi}{s}}$$ which holds for any $a\in \R$ and $s>0$ (see appendix; note that for a fixed second argument, the upper incomplete $\Gamma$-function is entire in the first argument, see \cite[pp. 899, 1032ff.]{gradshteyn2015table}).  Therefore, we get the integral

%formula 8.351.3, p. 
 %(see the wolfram functions website \url{http://functions.wolfram.com/06.06.26.0005.01}, it can also be found in the paper \url{https://arxiv.org/pdf/1507.05571}; , so we get the integral
\begin{align*}
= \xi^{a+1} \int_0^1 s^{c-1} (1-s)^{b-1} \MeijerG{2}{0}{1}{2}{1}{-a-1,0}{\frac{\xi}{s}}ds
\end{align*}
applying the inversion identity for Meijer's $G$-function (see \cite[p. 209, 5.3.1.(9))]{erdelyi1953higher})
\begin{align*}
=\xi^{a+1} \int_0^1 s^{c-1} (1-s)^{b-1} \MeijerG{0}{2}{2}{1}{2+a,1}{0}{\frac{s}{\xi}}ds
\end{align*}
This is the Euler transform of Meijer's $G$-function (see \cite[p. 214, 5.5.2.(5)]{erdelyi1953higher}) and (as the conditions $2+1<2(0+2)$ and $|\arg(\xi^{-1})| < \frac{\pi}{2}$ (as $\xi>0$) and $1-c-b<1-c$ (as $b>0$)  are satisfied) equals
\begin{align*}
=\xi^{a+1} \Gamma(b) \MeijerG{0}{3}{3}{2}{1-c,2+a,1}{0,1-c-b}{\xi^{-1}}
\end{align*}
Using again the inversion identity for Meijer's $G$-function
\begin{align*}
=\xi^{a+1} \Gamma(b) \MeijerG{3}{0}{2}{3}{1,b+c}{c,-1-a,0}{\xi}
\end{align*}
Plugging in $a= 6\alpha-k-3$, $b=2\alpha$, $c=3-4\alpha$ results in
\begin{align*}
=\xi^{6a-k-2} \Gamma(2\alpha) \MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}
\end{align*}
%(note that this agrees with the result in the old approach, after `bringing the power of $\xi$ into Meijer's $G$-function'.)

### 2018-10-21 introduction.tex
fixed errors in bib file and corresponding references in the LaTeX files

Hyperbolic random graphs were suggested as a suitable model for complex networks by Krioukov et al.~\cite{krioukov2010hyperbolic}, exhibiting the three main characteristics commonly found in complex networks: 
\begin{enumerate}[\upshape 1)]
\item Broad degree distribution (power-law/scale-free),
\item Strong clustering (community structure),
\item Small path lengths (small world phenomenon).
\end{enumerate}
Indeed, in their seminal paper they showed that these graphs have a power-law degree distribution and exhibit strong clustering, while Bringmann et al.~\cite{bringmann2016average} showed that hyperbolic random graphs have doubly logarithmic shortest path lengths. Many other results regarding hyperbolic random graphs have since been established: for instance the existence of efficient routing algorithms~\cite{bringmann2017greedy}, the size of the largest component~\cite{BodeFountoulakisMuller2015},~\cite{fountoulakis2018law}, [More?] and the number of cliques and the largest clique~\cite{blasius2018cliques}.

@@

%, which without loss of generality we take to be $-1$~[CITE]
%The \emph{Poincar\'{e} disc model} for the hyperbolic plane is defined by equipping the open unit disc in $\R^2$ with the metric given by the differential form $\dd s^2 = 4 \frac{\dd x^2 + \dd y^2}{(1 - x^2 - y^2)^2}$.
The hyperbolic plane is an infinite $2$-dimensional manifold with constant negative curvature. Many different representations of the hyperbolic plane exist and we refer to~\cite{Anderson1999} \cite{Katok1992} \cite{Beardon1983} for an introduction to hyperbolic geometry.  We will work with the \emph{native representation} of the hyperbolic plane which takes the elements of $\R^2$ in polar coordinates $(r,\theta)$ as the underlying set of points and where distances are determined by the hyperbolic law of cosines.

@@

\MS{Here, I would just say that $B(a,b)$ denotes the beta function $B^-(x,a,b)$ the lower incomplete beta function etc. and put/refer to the integral representations in the appendix, as already done for the Meijer-G function.}
%\subsubsection{Special functions}
We let $B(a,b)$ denote the beta-function 
%\[
%	B(a,b) = \int_0^1 t^{a-1}(1-t)^{b-1} \, dt
%\]
and $B^-(x,a,b)$ the lower incomplete beta-function
\[
	B^-(x,a,b) = \int_x^1 t^{a-1}(1-t)^{b-1} \, dt.
\]
In addition we write $\Gamma(z)$ for the Gamma function, $\Gamma^+(q,z)$ for the incomplete Gamma function
\[
	\Gamma^+(q,z) = \int_z^\infty t^{q - 1} e^{-t} \dd t
\]
and define $\Gamma^\ast(q,z) := \Gamma^+(1 + q, z) + \Gamma^+(q, z)$. Finally, we write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote Meijer's G-Function~\cite{meijer1946g}, see Appendix~\ref{sec:Meijer_G_functions} for more details.

>>>>

Hyperbolic random graphs were suggested as a suitable model for complex networks by Krioukov et al.~\cite{krioukov2010hyperbolic}, exhibiting the three main characteristics commonly found in complex networks: 
\begin{enumerate}[\upshape 1)]
\item Broad degree distribution (power-law/scale-free),
\item Strong clustering (community structure),
\item Small path lengths (small world phenomenon).
\end{enumerate}
Indeed, in their seminal paper they showed that these graphs have a power-law degree distribution and exhibit strong clustering, while Bringmann et al.~\cite{bringmann2016average} showed that hyperbolic random graphs have doubly logarithmic shortest path lengths. Many other results regarding hyperbolic random graphs have since been established: for instance the existence of efficient routing algorithms~\cite{bringmann2017greedy}, the size of the largest component~\cite{bode2015largest},~\cite{fountoulakis2018law}, [More?] and the number of cliques and the largest clique~\cite{blasius2018cliques}.

@@

%, which without loss of generality we take to be $-1$~[CITE]
%The \emph{Poincar\'{e} disc model} for the hyperbolic plane is defined by equipping the open unit disc in $\R^2$ with the metric given by the differential form $\dd s^2 = 4 \frac{\dd x^2 + \dd y^2}{(1 - x^2 - y^2)^2}$.
The hyperbolic plane is an infinite $2$-dimensional manifold with constant negative curvature. Many different representations of the hyperbolic plane exist and we refer to~\cite{anderson2006hyperbolic} \cite{katok1992fuchsian} \cite{beardon2012geometry} for an introduction to hyperbolic geometry.  We will work with the \emph{native representation} of the hyperbolic plane which takes the elements of $\R^2$ in polar coordinates $(r,\theta)$ as the underlying set of points and where distances are determined by the hyperbolic law of cosines.

@@

\MS{Here, I would just say that $B(a,b)$ denotes the beta function $B^-(x,a,b)$ the lower incomplete beta function etc. and put/refer to the integral representations in the appendix, as already done for the Meijer-G function.}
%\subsubsection{Special functions}
We let $B(a,b)$ denote the beta-function 
%\[
%	B(a,b) = \int_0^1 t^{a-1}(1-t)^{b-1} \, dt
%\]
and $B^-(x,a,b)$ the lower incomplete beta-function
\[
	B^-(x,a,b) = \int_x^1 t^{a-1}(1-t)^{b-1} \, dt.
\]
In addition we write $\Gamma(z)$ for the Gamma function, $\Gamma^+(q,z)$ for the incomplete Gamma function
\[
	\Gamma^+(q,z) = \int_z^\infty t^{q - 1} e^{-t} \dd t
\]
and define $\Gamma^\ast(q,z) := \Gamma^+(1 + q, z) + \Gamma^+(q, z)$. Finally, we write $U(a,b,z)$ for the hypergeometric U-function (also called Tricomi's confluent hypergeometric function), which for $a,b,z\in \mathbb{C}$, $b \not \in \mathbb{Z}_{\leq 0}$, $\mathrm{Re}(a), \mathrm{Re}(z) >0$ has the integral representation 
\[
	U(a,b,z) = \frac{1}{\Gamma(a)} \int_0^\infty e^{-zt} t^{a-1} (1+t)^{b-a-1} dt,
\] 
see~\cite[p.255 Equation (2)]{erdelyi1953higher}, and let $\MeijerGnew{m}{\ell}{p}{q}{{\bf a}}{{\bf b}}{z}$ denote Meijer's G-Function~\cite{meijer1946gfunction}, see Appendix~\ref{sec:Meijer_G_functions} for more details.
